A Study of Common Bug Fix Patterns in Rust \thanks{Competing Interests: We have no competing interests and are funded by a Discovery Grant from Canada's Natural Science and Engineering Research Council.}

### 2023-11-12 discussion.tex
final pass

We discuss three threats to external validity: selection bias, incorrect commit/bug information reported by developers, and changes to the Rust programming language. Selection bias is an issue because our benchmarks may differ systematically from the set of Rust programs in the world. Our selection of projects might be biased and therefore we might not have presented patterns that may exist in other projects. Another threat to external validity can be the developers reporting commits as bug fixing commits while they are not really fixing functionality. Similarly, a commit message might not contain our target keywords while the commit is associated with bug fixing changes. A third threat is that Rust is a relatively new programming language, and its syntax is prone to evolving over time. Changes in Rust's grammar, such as the introduction of new keywords like \verb+async/await+ in version 1.39, have the potential to introduce new bug categories or modify existing ones. Thus, one must consider language evolution as a potential external threat to the validity of our research findings; this threat can be mitigated by redoing our analysis in the future using our methodology with updated versions of our published artifacts.

>>>>

We discuss three threats to external validity: selection bias, incorrect commit/bug information reported by developers, and changes to the Rust programming language. (1) Selection bias is an issue because our benchmarks may differ systematically from the set of Rust programs in the world. Our selection of projects might be biased and therefore we might not have presented patterns that may exist in other projects. We aimed to mitigate this threat by choosing all of the most-starred projects, but this biases towards popular projects by definition. (2) Another threat to external validity can be the developers reporting commits as bug fixing commits while they are not really fixing functionality. Similarly, a commit message might not contain our target keywords while the commit is associated with bug fixing changes. This is simply an underlying assumption of our work and may cause us to miss some bugs; we do not believe it should be a systematic threat, and is shared by much other work on understanding bug fixes. (3) A third threat is that Rust is a relatively new programming language, and its syntax is prone to evolving over time. Changes in Rust's grammar, such as the introduction of new keywords like \verb+async/await+ in version 1.39, have the potential to introduce new bug categories or modify existing ones. Thus, one must consider language evolution as a potential external threat to the validity of our research findings; this threat can be mitigated by redoing our analysis in the future using our methodology with updated versions of our published artifacts.

### 2023-11-12 methodology.tex
final pass

In this section, we provide an overview of our methodology, providing full details in later subsections. Ruxanne uses a pipeline for finding classes of pervasive patterns in Rust programs. This pipeline parses program versions, computes differences between them, embeds them in fixed-size datapoints, and then clusters these datapoints. After conducting a manual analysis on the obtained clusters, we refine them to obtain our proposed bug fix patterns.

@@

Now that we have set up our code analysis pipeline, we can start mining the Rust repositories. We target the top 18 most-starred Rust projects on GitHub, at the time of data collection. Table~\ref{table:repos} summarizes our benchmarks. Because we are interested in changes, we computed (using process metrics provided by our mining library) the average code churn of the files within the projects. swc-project/swc, nushell/nushell, and denoland/deno are the projects with largest number of LOCs. Unsurprisingly, we captured a lot of instances from these projects in our final clusters.

>>>>

In this section, we provide an overview of our methodology, with full details to appear in later subsections. Ruxanne uses a pipeline for finding classes of pervasive patterns in Rust programs. This pipeline parses program versions, computes differences between them, embeds them in fixed-size datapoints, and then clusters these datapoints. After conducting a manual analysis on the obtained clusters, we refine them to obtain our proposed bug fix patterns.

@@

Now that we have set up our code analysis pipeline, we can start mining the Rust repositories. We target the top 18 most-starred Rust projects on GitHub, at the time of data collection. Table~\ref{table:repos} summarizes our benchmarks. Because we are interested in changes, we computed (using process metrics provided by our mining library) the average code churn of the files within the projects. swc-project/swc, nushell/nushell, and denoland/deno are the projects with the most LOCs. Unsurprisingly, we captured a lot of instances from these projects in our final clusters.

### 2023-11-12 related.tex
final pass

\cite{pan2009toward} used a bug fix pattern extractor tool to automatically parse and detect bug patterns within bug fix hunks with fewer than seven statements. They ignored larger bug fix hunks---such hunks tended to exhibit random changes rather than having patterns that they could derive meaning from. Though they analyzed Java projects, their reported patterns were not specific to Java. \cite{martinez2015mining,martinez2012mining} exploited the frequency of observed patterns to introduce a heuristic patch searching method. In their empirical evaluation, they concluded that choosing repair actions probabilistically (weighted by frequency) can help with reducing the search space, hence creating a more effective repair tool.

@@

\paragraph{Automated Patch Correctness Assessment.}
The final step in automated program repair is to validate the proposed changes. One substantial challenge is overfitting, where synthesized patches may pass test cases but fail to be correct. Automated patch correctness assessment (APCA) aims to inspect the correctness of generated patches using dynamic and static analyses. While our project has a different objective than APCA, we share similarities in design decisions with APCA systems, particularly in feature extraction. \cite{ye2021automated} propose ODS, an overfitting patch detection system, which assumes that code features capturing universal correctness properties can be utilized to classify overfitting patches across program repair systems and software projects. ODS extracts 202 manually crafted code features by comparing fixed and buggy code, and employs supervised learning to build a distributional model for classifying unseen patches, achieving over 70\% accuracy in evaluations on Defects4J, Bugs.jar, and Bears benchmarks. Similarly, our feature extraction involves manual crafting of features. In another study, \cite{tian2022change} present an enhanced APCA model that transforms the problem into a question-answering scenario and incorporates a neural architecture to learn semantic correlations between bug reports and commit messages. Additionally, \cite{lin2022context} introduce \textsc{Cache}, a neural-based context-aware AST embedding method that captures both the changed part and the unchanged part (context code) of patches. Evaluation results demonstrate that context knowledge significantly enhances the performance and accuracy of systems that attempt to automatically assess patch correctness (which is not our goal).

@@

Our work is more like that of~\cite{hanam2016discovering}. However, our novel program embedding approach differs from theirs on certain key decisions. In~\cite{hanam2016discovering}, they introduced the Feature Properties table, which is a categorization for program elements. Embedding programs based on this table resulted in an order-sensitive embedding. However, it made the datapoints sparse; our datapoints are more dense than theirs. More specifically, the number of our dimensions is more than 11$\times$ smaller.

>>>>

\cite{pan2009toward} used a bug fix pattern extractor tool to automatically parse and detect bug patterns within bug fix hunks with fewer than seven statements. They ignored larger bug fix hunks---such hunks tended to exhibit random changes rather than having patterns that they could derive meaning from. Though they analyzed Java projects, their reported patterns were not specific to Java, in the same way that our general patterns are not specific to Rust. \cite{martinez2015mining,martinez2012mining} exploited the frequency of observed patterns to introduce a heuristic patch searching method. In their empirical evaluation, they concluded that choosing repair actions probabilistically (weighted by frequency) can help with reducing the search space, hence creating a more effective repair tool.

@@

\paragraph{Automated Patch Correctness Assessment.}
The final step in automated program repair is to validate the proposed changes. One substantial challenge is overfitting, where synthesized patches may pass test cases but fail to be correct. Automated patch correctness assessment (APCA) aims to inspect the correctness of generated patches using dynamic and static analyses. While our project has a different objective than APCA, we share similarities in design decisions with APCA systems, particularly in feature extraction. \cite{ye2021automated} propose ODS, an overfitting patch detection system, which assumes that code features capturing universal correctness properties can be utilized to classify overfitting patches across program repair systems and software projects. ODS extracts 202 manually crafted code features by comparing fixed and buggy code, and employs supervised learning to build a distributional model for classifying unseen patches, achieving over 70\% accuracy in evaluations on Defects4J, Bugs.jar, and Bears suites. Similarly, our feature extraction involves manual crafting of features. In another study, \cite{tian2022change} present an enhanced APCA model that transforms the problem into a question-answering scenario and incorporates a neural architecture to learn semantic correlations between bug reports and commit messages. Additionally, \cite{lin2022context} introduce \textsc{Cache}, a neural-based context-aware AST embedding method that captures both the changed part and the unchanged part (context code) of patches. Evaluation results demonstrate that context knowledge significantly enhances the performance and accuracy of systems that attempt to automatically assess patch correctness (which is not our goal).

@@

Our work is more like that of~\cite{hanam2016discovering}. However, our novel program embedding approach differs from theirs on certain key decisions. Their work introduced the Feature Properties table, which is a categorization for program elements. Embedding programs based on this table resulted in an order-sensitive embedding. However, it made the datapoints sparse; our datapoints are more dense than theirs. Specifically, we have over 11$\times$ fewer dimensions than they do.

### 2023-11-12 results.tex
final pass

\noindent\textbf{Description:} Rust is a null-safe language, meaning that object references cannot take on null values. If a developer wants a variable to contain the equivalent of null, they would have to use the \texttt{Option} type. \texttt{Option} is essentially an \texttt{enum} with two variants: \texttt{Some}, which indicates that a variable has a value; and \texttt{None}, which indicates the absence of any value. In this change pattern, the developer changes the type of a struct field to an \texttt{Option} of that type, hence enabling them to store no value in that field. 
For instance, in the Rust project swc-project/swc (a fast TypeScript/JavaScript compiler), the developer needed to modify the CSS parser source code to account for empty @media queries\footnote{\scriptsize \url{https://github.com/swc-project/swc/commit/75a14f98b7370226115ee24eec6eb8c802bd4837}}. As shown in the snippet above, they changed type \texttt{MediaQueryList} to \verb+Option<MediaQueryList>+, and modified the other parts of the source code accordingly. \\

@@

In Rust, the match control flow construct is used for comparing a value against multiple patterns, and executing the code associated with the first pattern that matches with our value. A pattern and its associated code are called a match arm. Match statements can be similar to switch statements in other languages, like C, although match statements in Rust are able to match more complicated patterns.\\

@@

\noindent\textbf{Description:} In this pattern, a change occurs in an arm's pattern. Like the code snippet provided above, it can be a change in a match pattern's type. It also can be matching to a different enum variant or string. The fix pattern can happen both for refactoring or debugging purposes. \\

@@

\noindent\textbf{Description:} This pattern is referred to as redundant clone in Clippy Lints. Clippy\footnote{\url{https://github.com/rust-lang/rust-clippy}} is a linter tool for Rust that catches common mistakes. There are more than 500 lints included in Clippy. Since detecting whether a clone is redundant is undecidable, Clippy conservatively estimates redundancy. By definition, removing a clone that is indeed redundant does not modify functional aspects of the program behaviour. (Removing a non-redundant clone introduces a compile error or, worse yet, a bug.) However, such a removal can greatly improve the nonfunctional property of performance, if the program was formerly cloning large objects. \\

@@

Like the previous pattern, this change does not affect the program behaviour; it is done for performance purposes. In one of the commits\footnote{\scriptsize \url{https://github.com/tauri-apps/tauri/commit/a280ee90af0749ce18d6d0b00939b06473717bc9}} of the project tauri-apps/tauri, changing a clone to borrowing significantly improved CPU usage. In the commit history, developers discussed how the cloning of types such as HashMap was expensive. \\

>>>>

\noindent\textbf{Description:} Rust is a null-safe language, meaning that object references cannot take on null values. If a developer wants a variable to contain the equivalent of null, they must use the \texttt{Option} type. \texttt{Option} is essentially an \texttt{enum} with two variants: \texttt{Some}, which indicates that a variable has a value; and \texttt{None}, which indicates the absence of any value. In this change pattern, the developer changes the type of a struct field to an \texttt{Option} of that type, hence enabling them to store no value in that field. 
For instance, in the Rust project swc-project/swc (a fast TypeScript/JavaScript compiler), the developer needed to modify the CSS parser source code to account for empty @media queries\footnote{\scriptsize \url{https://github.com/swc-project/swc/commit/75a14f98b7370226115ee24eec6eb8c802bd4837}}. As shown in the snippet above, they changed type \texttt{MediaQueryList} to \verb+Option<MediaQueryList>+, and modified the other parts of the source code accordingly. \\

@@

In Rust, the match control flow construct is used for comparing a value against multiple patterns, and executing the code associated with the first pattern that matches. A pattern and its associated code are called a match arm. Match statements can be similar to switch statements in other languages, like C, although match statements in Rust are able to match more complicated patterns.\\

@@

\noindent\textbf{Description:} In this pattern, a change occurs in an arm's pattern. Like the code snippet provided above, it can be a change in a match pattern's type. It also can be matching to a different enum variant or string. The fix pattern can happen both for refactoring or bug-fixing purposes. \\

@@

\noindent\textbf{Description:} This pattern is referred to as redundant clone in Clippy Lints. Clippy\footnote{\url{https://github.com/rust-lang/rust-clippy}} is a linter tool for Rust that catches common mistakes. There are more than 500 lints included in Clippy. Since detecting whether a clone is redundant is undecidable, Clippy conservatively estimates redundancy. By definition, removing a clone that is indeed redundant does not modify functional aspects of the program behaviour. (Removing a non-redundant clone introduces a compile error or, worse yet, a bug.) However, a removal of a redundant clone can greatly improve the nonfunctional property of performance, if the program was formerly cloning large objects. \\

@@

Like the previous pattern, this change does not affect the program behaviour; it is done for performance purposes. In one of the commits\footnote{\scriptsize \url{https://github.com/tauri-apps/tauri/commit/a280ee90af0749ce18d6d0b00939b06473717bc9}} of the project tauri-apps/tauri, changing a clone to borrowing significantly improved CPU usage: in the commit history, developers discussed how the cloning of types such as HashMap was expensive. \\
